%\documentclass[11pt]{article}
\documentclass[preprintnumbers,prd,superscriptaddress,notitlepage,nofootinbib] {revtex4-1}

\usepackage{geometry, amsmath, amsthm, latexsym, amssymb, graphicx}
\geometry{margin=1in, headsep=0.25in}
\usepackage[usenames, dvipsnames]{color}

%comments by co-authors
\newcommand{\AFR}[1]{{\color{red}AFR: #1}}

%journals
\newcommand{\mnras}{{\em Mon. Not. Roy. Astron. Soc. }}
%\newcommand{\apj}{{\em Astrophys. Jour. }}
%\newcommand{\apjl}{{\em Astrophys. J. Let. }}
\newcommand{\apjs}{ApJS}
\newcommand{\jcap}{JCAP}
%\newcommand{\physrep}{{\em Phys. Rept. }}
\newcommand{\aap}{{\em Astron. Astrophys. }}
%\newcommand{\prd}{{\em Phys. Rev.}}
%\newcommand{\aj}{AJ }

%others
\newcommand{\lya}{Ly-$\alpha$}
\newcommand{\kms}{{\rm km~s^{-1}}}  
\newcommand{\ikms}{{\rm s~km^{-1}}}
\newcommand{\iMpc}{{\rm Mpc^{-1}}}
\newcommand{\kpar}{k_\parallel}
\newcommand{\Plin}{P_{\rm lin}}

\begin{document}

\title{Parameterization of the \lya\ forest power spectrum}

\author{Andreu Font-Ribera \footnote{author list alphabetized}}
\email{a.font@ucl.ac.uk} 
\affiliation{Department of Physics and Astronomy, University College London, 
Gower Street, London, United Kingdom}
\author{Patrick McDonald}
\email{PVMcDonald@lbl.gov} 
\affiliation{Lawrence Berkeley National Laboratory, One Cyclotron Road,
Berkeley, CA 94720, USA}
\author{An\v{z}e Slosar}
\email{anze@bnl.gov} 
\affiliation{Brookhaven National Laboratory, Upton, NY 11973, USA}

\date{\today}

\begin{abstract}
Discuss unique role of \lya\ forest power spectrum, running and neutrino mass.
Discuss the need of emulators and expensive hydrodynamical simulations, and
the critical choice of likelihood parameterization.
Mention BOSS, eBOSS, DESI and high-resolution spectra.
\end{abstract}

\maketitle

\input{Introduction}

\input{Overview}

\input{Linear}

\input{Simulations}

\input{Density}

\input{Lya}

\input{Discussion}

\section*{Acknowledgements}
AFR acknowledges support by an STFC Ernest Rutherford Fellowship, grant reference ST/N003853/1.
AS acknowledges hospitality of the University College London.
This work was partially enabled by funding from the UCL Cosmoparticle
Initiative.

\bibliography{main}
\bibliographystyle{JHEP}

%\appendix


\end{document}
